### 2018-05-23 Paper/examples.tex
..

As discussed earlier in this section, the reduced order PCE's are verified for
predictive accuracy in a least-squares sense and a probabilistic sense.
Estimates for $\epsilon_{\mbox{\tiny{L-2}}}$ based on 50 samples and
corresponding function evaluations in the original 7D parameter space were
found to be 0.0551 and 0.0112 for the 4D and 5D PCE's respectively. In other
words, the 4D PCE is accurate within 5.52$\%$ and the 5D PCE is accurate within
1.12$\%$ of predictions based on the borehole function. We note  that although
the convergence in the case of a 5D PCE is slower (Figure~\ref{fig:conv_bore}),
its predictive accuracy is higher than the 4D PCE. This illustrates the
tradeoff between accuracy and computational efficiency for the present 
problem. Generally, the required level of accuracy is problem dependent. 
The present framework allows for moving towards higher fidelity 
reduced order surrogates based on the ranking of the parameter sensitivities.

@@

Figure~\ref{fig:conv_bore}~(right) illustrates a comparison of the PDFs
of the
discharge, $\mathcal{Q}$ obtained by propagating 10$^6$ random
samples through the 7D PCE in the original input parameter domain as well as
the reduced order PCEs constructed in 4 and 5 dimensions. 
It is evident from this plot, that the PDFs agree quite favorably with each
other with respect to the modal estimate as well as the uncertainty associated with 
$\mathcal{Q}$. 
Consequently, it can be said that the reduced order PCE is verified in a
probabilistic sense. In other words, the mode as well as the uncertainty in the
observable is reliably captured and predicted by the reduced order PCE. 
% 
\subsection{Nonlinear Oscillator}
We consider an undamped nonlinear oscillator with only one degree of
freedom as illustrated in Figure~\ref{fig:osc}. 
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.5\textwidth]{./Figures/oscillator}
\caption{Schematic illustration of an undamped non-linear oscillator system with one degree of freedom.}
\label{fig:osc}
\end{center}
\end{figure}
This system has been studied extensively for its reliability using
the following state function~\cite{Bucher:1989,Bucher:1990,Rajashekhar:1993,Schueremans:2005,Gayton:2003}:
%
\be
g(\bm{X}) = 3r - \left|z_\text{max}\right| = 3r - 
\left|\frac{2F}{m\omega_0^2}\sin\left(\frac{\omega_0t_1}{2}\right)\right|,
\label{eq:limit}
\ee
\noindent where $z_\text{max}$ is the maximum displacement response of the system, $r$ denotes the displacement at
which either spring yields, $\bm{X}=(m,k_1,k_2,r,F,t_1)^T$ is the vector 
of uncertain parameters, and   
$\omega_0~=~\sqrt{(k_1+k_2)/m}$. The uncertain parameters are considered to be normally distributed with mean
and variance as provided below in Table~\ref{tab:osc}.

@@

To verify the accuracy of the third-order PCE, we estimate the relative L-2
norm of the difference between predictions of  $g(\bm{X})$, obtained using the
actual function in \eqref{eq:limit} and the 
PCE using \eqref{eq:l2}, and a validation test suite. The test suite comprised
model evaluations at 100 independent Monte-Carlo samples in the full 
parameter space. The value of $\epsilon_{\mbox{\tiny{L-2}}}$ 
was found to be 0.0835 i.e. the third-order PCE is accurate within 8.36$\%$.
Furthermore, the accuracy is assessed in a probabilistic sense by comparing
PDFs of $g(\bm{X})$, obtained using the function and the 3D PCE as illustrated
in Figure~\ref{fig:conv_osc}~(right).  Clearly, the two PDFs are in excellent
agreement with each other. Hence, the ROS constructed using the methodology
presented in section~\ref{sec:method} could be used reliably to 
quantify the uncertainty in $g(\bm{X})$.
% due to parametric uncertainty
%associated with its functional description in~\eqref{eq:limit}, efficiently.

@@

%\begin{table}[htbp]
%\begin{tabular}{cc}
%
%  \includegraphics[width=0.45\textwidth]{./Figures/u_soln.png} &
%\begin{tabular}{cl}
%\toprule
%$\theta_i$ & \textbf{Distribution} \\ 
%\bottomrule
%$\kappa$ & $\mathcal{U}$[0.05,0.1] \\
%$c$ & $\mathcal{U}$[1.0,2.0] \\
%$\alpha_i$ & $\mathcal{U}$[0.0,4.0] \\
%\bottomrule
%\end{tabular}
%\\
%\end{tabular}
%\end{table}

We aim to construct a reduced-order surrogate for the following QoI:
\be
\mathcal{F}(\kappa, c, \theta) = \frac{1}{|D|} \int_D u(x; \kappa, c, \theta) \, dx, 
\label{eq:qoi}
\ee
%
where $D$ is the region $[2/5, 3/5] \times [2/5, 3/5] \subset \Omega$, 
and $|D|$ denotes the area of $D$. 
While this model is considerably more complex than the previous numerical
examples, it can still be solved efficiently.
The equation was discretized using finite-differences and Newton's method
was used to solve the resulting nonlinear system.~\alennote{Give the
number of degrees of freedom.} 
We computed converged estimates of the Sobol
total-effect index $\mathcal{T}(\theta_i)$, reported 
in Figure~\ref{fig:sense_elliptic}~(left).~\alennote{Say how many PDE solves
were used.}
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semilinear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
%
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$,
plotted in Figure~\ref{fig:sense_elliptic}~(right), are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. As $N$
is increased from 5 to 20, estimates of the screening metric are observed to converge.
Based on the trends observed in Figure~\ref{fig:sense_elliptic}, it can be said that
the uncertainty in the QoI in \eqref{eq:qoi} is largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an ROS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic} (left). 
As expected, the ROS converges considerably faster. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) versus
$\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the ROS in the present test problem
is expected to be much smaller. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_elliptic}
\caption{Left: Logarithm of $\epsilon_{\mbox{\tiny{LOO}}}$ is plotted against sample size for 
PCEs constructed in 10 and 5 dimensions to compare their convergence characteristics. 
Right: PDF of the QoI, $\mathcal{F}(\kappa, c, \theta)$ in~\eqref{eq:semilinear} is
plotted using the full-surrogate and the reduced-order PC surrogate.}
\label{fig:conv_elliptic}
\end{center}
\end{figure}

Once again, we verify the accuracy of the ROS by estimating $\epsilon_{\mbox{\tiny{L-2}}}$
using model evaluations at 100 independent Monte-Carlo samples in the 10D parameter
space, and the ROS was found to be accurate
within 5$\%$. 
%The L-2 norm was estimated to be 0.05 i.e. 
Additionally, we compare PDFs of the QoI, $\mathcal{F}(\kappa,c,\theta)$,
generated using the full-space and the reduced-space surrogates in
Figure~\ref{fig:conv_elliptic}~(right). The two PDFs are observed to be in
excellent agrement with each other. Hence, the ROS is successfully verified and
could be used in lieu of the full-space surrogate to reduce computational
effort. 
\alennote{Regarding the PDFs reported in figures 3, 6, 9: I think since we 
are sampling the model at a large number of points to compute the
total Sobol indices, we might as well save those sames, and then
use them to compute a KDE of the ``true'' PDF; that way, each of these
PDF comparison plots will report: true, FSS, ROS PDFs. I think that will
make the case 
for the presented method stronger, given we remain close to the true PDF.}

>>>>

As discussed earlier in this section, the reduced order PCE's are verified for
predictive accuracy in a least-squares sense and a probabilistic sense.
Estimates for $\epsilon_{\mbox{\tiny{L-2}}}$ based on 50 samples in the
validation test suite were
found to be 0.0551 and 0.0112 for the 4D and 5D PCE's respectively. In other
words, the 4D PCE is accurate within 5.52$\%$ and the 5D PCE is accurate within
1.12$\%$ of predictions based on the borehole function. We note  that although
the convergence in the case of a 5D PCE is slower (Figure~\ref{fig:conv_bore}),
its predictive accuracy is higher than the 4D PCE. This illustrates the
tradeoff between accuracy and computational efficiency for the present 
problem. Generally, the required level of accuracy is problem dependent. 
The present framework allows for moving towards higher fidelity 
reduced order surrogates based on the ranking of the parameter sensitivities.

@@

Figure~\ref{fig:conv_bore}~(right) illustrates a comparison of the PDFs
of the
discharge, $\mathcal{Q}$ obtained by propagating 10$^6$ random
samples through the 7D PCE in the original input parameter domain as well as
the reduced order PCEs constructed in 4 and 5 dimensions. A normalized histogram
plot using model evaluations in the validation test suite is also included.
It is evident from this plot, that the PDFs agree quite favorably with each
other as well as the model-based histogram with respect to the modal estimate
as well as the uncertainty associated with $\mathcal{Q}$. 
Consequently, it can be said that the reduced order PCE is verified in a
probabilistic sense. In other words, the mode as well as the uncertainty in the
observable is reliably captured and predicted by the reduced order PCE. 
% 
\subsection{Nonlinear Oscillator}
We consider an undamped nonlinear oscillator with only one degree of
freedom as illustrated in Figure~\ref{fig:osc}. 
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.5\textwidth]{./Figures/oscillator}
\caption{Schematic illustration of an undamped non-linear oscillator system with one degree of freedom.}
\label{fig:osc}
\end{center}
\end{figure}
This system has been studied extensively for its reliability using
the following state function~\cite{Bucher:1989,Bucher:1990,Rajashekhar:1993,Schueremans:2005,Gayton:2003}:
%
\be
g(\bm{X}) = 3r - \left|z_\text{max}\right| = 3r - 
\left|\frac{2F}{m\omega_0^2}\sin\left(\frac{\omega_0t_1}{2}\right)\right|,
\label{eq:limit}
\ee
\noindent where $z_\text{max}$ is the maximum displacement response of the system, $r$ denotes the displacement at
which either spring yields, $\bm{X}=(m,k_1,k_2,r,F,t_1)^T$ is the vector 
of uncertain parameters, and   
$\omega_0~=~\sqrt{(k_1+k_2)/m}$. The uncertain parameters are considered to be normally distributed with mean
and variance as provided below in Table~\ref{tab:osc}.

@@

To verify the accuracy of the third-order PCE, we estimate the relative L-2
norm of the difference between predictions of  $g(\bm{X})$, obtained using the
actual function in \eqref{eq:limit} and the 
PCE using \eqref{eq:l2}, and a validation test suite. The test suite comprised
model evaluations at 100 independent Monte-Carlo samples in the full 
parameter space. The value of $\epsilon_{\mbox{\tiny{L-2}}}$ 
was found to be 0.0835 i.e. the third-order PCE is accurate within 8.36$\%$.
Furthermore, the accuracy is assessed in a probabilistic sense by comparing
in Figure~\ref{fig:conv_osc}~(right),
PDFs of $g(\bm{X})$, obtained using the full-surrogate, reduced-order 
surrogates constructed 3 and 4 dimensions, and a normalized histogram based
on model evaluation in the validation test suite. The three PDFs are observed
to be in excellent agreement with each other as well as the model-based
histogram. Hence, the ROS constructed using the methodology
presented in section~\ref{sec:method} could be used reliably to 
quantify the uncertainty in $g(\bm{X})$ with reduced effort.

@@

We aim to construct a reduced-order surrogate for the following QoI:
\be
\mathcal{F}(\kappa, c, \theta) = \frac{1}{|D|} \int_D u(x; \kappa, c, \theta) \, dx, 
\label{eq:qoi}
\ee
%
where $D$ is the region $[2/5, 3/5] \times [2/5, 3/5] \subset \Omega$, 
and $|D|$ denotes the area of $D$. 
While this model is considerably more complex than the previous numerical
examples, it can still be solved efficiently.
The equation was discretized using finite-differences, and Newton's method
was used to solve the resulting nonlinear system on a (100$\times$100) 2D
cartesian grid.~\alennote{Give the
number of degrees of freedom.} 
We computed converged estimates of the Sobol
total-effect index $\mathcal{T}(\theta_i)$, reported 
in Figure~\ref{fig:sense_elliptic}~(left) by solving the above non-linear
system at 500 samples in the 10D parameter space. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semilinear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
%
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$,
plotted in Figure~\ref{fig:sense_elliptic}~(right), are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. As $N$
is increased from 5 to 20, estimates of the screening metric are observed to converge.
Based on the trends observed in Figure~\ref{fig:sense_elliptic}, it can be said that
the uncertainty in the QoI in \eqref{eq:qoi} is largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an ROS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic} (left). 
As expected, the ROS converges considerably faster. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) versus
$\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the ROS in the present test problem
is expected to be much smaller. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_elliptic}
\caption{Left: Logarithm of $\epsilon_{\mbox{\tiny{LOO}}}$ is plotted against sample size for 
PCEs constructed in 10 and 5 dimensions to compare their convergence characteristics. 
Right: PDF of the QoI, $\mathcal{F}(\kappa, c, \theta)$ in~\eqref{eq:semilinear} is
plotted using the full-surrogate and the reduced-order PC surrogate.}
\label{fig:conv_elliptic}
\end{center}
\end{figure}

Once again, we verify the accuracy of the ROS by estimating $\epsilon_{\mbox{\tiny{L-2}}}$
using model evaluations at 100 independent Monte-Carlo samples in the 10D parameter
space. The ROS was found to be accurate
within 5$\%$. In order to bolster confidence in the ROS, we compare PDFs of the QoI
as well as a normalized histogram plot based on sparse model evaluations in the  
validation test-suite, in 
Figure~\ref{fig:conv_elliptic}~(right). While the two PDFs are in favorable agreement,
the modal estimate and the spread in the QoI based on the histogram is also captured
by them. Hence, the ROS could be used with a reasonable degree of confidence to
quantify the uncertainty in $\mathcal{F}(\kappa, c, \theta)$ thereby leading to
a computational advantage in this case.
